\bfseries\LARGE\singlespacing{Transmission of respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-10-16 doc/paper/manuscript-lancet.tex

We analyzed a total of 448~saliva samples. We detected 15~influenza~B~(Flu B), 15~rhinoviruses~(HRV), 14~adenovirus (AdV), 3~SARS-CoV-2~(CoV), 2~metapneumoviruses~(MPV), and 1~parainfluenza~virus~(PIV) (\cref{fig:molecular-descriptives}a). The distribution of positive saliva samples varied between classes (\cref{fig:molecular-descriptives}b). For example, during the first three study weeks, all but one sample was positive for AdV in class~A, while the vast majority of positive Flu~B samples were found in class~B. To illustrate possible transmission chains within classes, we linked positive saliva samples of the same virus that were less than one week apart. Based on this spatiotemporal analysis, we identified 10~possible transmission chains (\cref{fig:molecular-network}). The longest transmission chains possibly occurred in January and refer to a cluster of AdV infections in class~A and Flu~B infections in class~B.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{../../results/mol-data/descriptives.pdf}
    \caption{\textbf{Molecular detection of respiratory viruses and transmission network based on spatiotemporal analysis of students' saliva samples}. \textbf{(a)}~Number of positive saliva samples by virus. \textbf{(b)}~Daily number of positive saliva samples (colored circles) and possible transmission chains within classes (directed arrows). Positive samples are linked if they belong to the same virus and are less than 1~week apart. Positive samples from the air and filters as blank squares aligned. Flu B: Influenza B, HRV: rhinovirus, AdV: adenovirus, CoV: SARS-CoV-2, MPV: metapneumovirus, PIV: parainfluenza virus.}
    \label{fig:molecular-descriptives}
\end{figure}

@@

Slightly fewer coughs were detected in audio recordings with air cleaners (\cref{fig:coughing}a). On average, there were 2.5~coughs/min with and 3.1~coughs/min without air cleaners ($\Delta=-$0.6~coughs/min; adjusted relative risk 0.93, 95\%-CrI 0.85 to 1.02). Significantly more coughs were detected in the room of class~B (adjusted relative risk 2.16, 95\%-CrI 1.90 to 2.46), suggesting an association with the observed variation in virus-specific transmission between classes. Indeed, cough was more frequent with Flu~B positivity (relative risk 1.25, 95\%-CrI 1.03 to 1.53) and slightly less frequent with AdV positivity (0.87, 95\%-CrI 0.49 to 1.12) (\cref{fig:coughing-association}b).

\begin{figure}[!htpb]
\centering
    \includegraphics{../../results/mol-data/saliva-by-condition.pdf}
    \includegraphics{../../results/env-data/estimation-results-figure.pdf}
    \includegraphics{../../results/epi-data/cases_by_condition.pdf}
    \includegraphics{results/cough-data/coughs-frequency-by-condition.pdf}
    \caption{\textbf{Risk of infection by study condition}. \textbf{(a)}~Number of positive saliva samples. \textbf{(b)}~Daily average aerosol number concentrations (CN in 1/cm$^3$) as boxplots. \textbf{(c)}~Number of respiratory cases. \textbf{(d)}~Daily average number of detected coughs per minute.}
    \label{fig:palas-results}
\end{figure}

>>>>

We analyzed a total of 448~saliva samples. We detected 15~influenza~B~(Flu B), 15~rhinoviruses~(HRV), 14~adenovirus (AdV), 3~SARS-CoV-2~(CoV), 2~metapneumoviruses~(MPV), and 1~parainfluenza~virus~(PIV) (\cref{fig:molecular-descriptives}a). The distribution of positive saliva samples varied between classes. For example, during the first three study weeks, all but one sample was positive for AdV in class~A, while the vast majority of positive Flu~B samples were found in class~B. To illustrate possible transmission chains within classes, we linked positive saliva samples of the same virus that were less than one week apart. Based on this spatiotemporal analysis, we identified 10~possible transmission chains (\cref{fig:molecular-descriptives}b). The longest transmission chains possibly occurred in January and refer to a cluster of AdV infections in class~A and Flu~B infections in class~B.

@@

\begin{figure}[!htpb]
    \centering
    \includegraphics{../../results/mol-data/saliva-distribution.pdf}
    \includegraphics{../../results/mol-data/network-plot.pdf}
    \caption{\textbf{Molecular detection of respiratory viruses and transmission network based on spatiotemporal analysis of students' saliva samples}. \textbf{(a)}~Number of positive saliva samples by virus. \textbf{(b)}~Daily number of positive saliva samples (colored circles) and possible transmission chains within classes (directed arrows). Positive samples are linked if they belong to the same virus and are less than 1~week apart. Positive samples from the air and filters as blank squares aligned. Flu B: Influenza B, HRV: rhinovirus, AdV: adenovirus, CoV: SARS-CoV-2, MPV: metapneumovirus, PIV: parainfluenza virus.}
    \label{fig:molecular-descriptives}
\end{figure}

@@

Slightly fewer coughs were detected in audio recordings with air cleaners (\cref{fig:infection-risk}d). On average, there were 2.5~coughs/min with and 3.1~coughs/min without air cleaners ($\Delta=-$0.6~coughs/min; adjusted relative risk 0.93, 95\%-CrI 0.85 to 1.02). Significantly more coughs were detected in the room of class~B (adjusted relative risk 2.16, 95\%-CrI 1.90 to 2.46), suggesting an association with the observed variation in virus-specific transmission between classes. Indeed, cough was more frequent with Flu~B positivity (relative risk 1.25, 95\%-CrI 1.03 to 1.53) and slightly less frequent with AdV positivity (0.87, 95\%-CrI 0.49 to 1.12) (\supp~figure~\zref{fig:coughing-association}b).

\begin{figure}[!htpb]
\centering
    \includegraphics{../../results/mol-data/saliva-by-study-condition.pdf}
    \includegraphics{../../results/env-data/aerosol-number-boxplot.pdf}
    \includegraphics{../../results/epi-data/cases_by_condition.pdf}
    \includegraphics{../../results/cough-data/coughs-frequency-by-condition.pdf}
    \caption{\textbf{Risk of infection by study condition}. \textbf{(a)}~Number of positive saliva samples. \textbf{(b)}~Daily average aerosol number concentrations as boxplots. \textbf{(c)}~Number of respiratory cases. \textbf{(d)}~Daily average number of detected coughs per minute.}
    \label{fig:infection-risk}
\end{figure}
